---
title: "Estudio de variables que afectan a la nota media obtenido por los alumnos"
author: "Jose Luis Alonso"
date: "2023-03-09"
output: html_document
---
## Load libreries
```{r}
library(readr);library(dplyr);library(ggplot2);library(plotly);library(shiny)
library(dplyr);library(plotly);library(tidyr);library(fmsb);library(radarchart)
library(ggthemes);library(hrbrthemes);library(ggridges);library(ggcorrplot);hrbrthemes::import_roboto_condensed()
library(hexbin)
```


#Load data
```{r}
data <- read_csv("../data/dataset.csv")
```


###¿Influye que un alumno sea desplazado o no en la nota?
#Calculo de las notas medias
```{r}
data$MeanGrade <- (data$`Curricular units 1st sem (approved)` + data$`Curricular units 2nd sem (approved)`) / 2
```


#Create a dataframe with only the columns of interest
```{r}
notas_desplazado <- data[,c("MeanGrade", "Displaced")]
```


#Create histogram
```{r}
ggplot(notas_desplazado, aes(x = MeanGrade, fill = as.factor(Displaced))) +
  geom_histogram(alpha = 0.5, position = position_stack(vjust = 0), 
                 color = "black", binwidth = 0.5, boundary = 0) +
  xlab("Media de notas obtenida") +
  ylab("Frecuencia") +
  ggtitle("Distribución de la media de notas obtenida por alumnos desplazados y no desplazados") +
  scale_fill_manual(values = c("red", "blue"), labels = c("Desplazado", "No desplazado"))
```


#Create density plot
```{r}
ggplot(notas_desplazado, aes(x=MeanGrade, fill=as.factor(Displaced))) +
  geom_density(alpha=0.5) +
  xlab("Media de notas obtenida") +
  ylab("Densidad") +
  ggtitle("Distribución de la media de notas obtenida por alumnos desplazados y no desplazados") +
  scale_fill_manual(values=c("red", "blue"), labels=c("Desplazado", "No desplazado"))
```


#We calculate students with and without scholarships separately.
```{r}
scholarship <- subset(data, data$`Scholarship holder` == 1)
no_scholarship <- subset(data, data$`Scholarship holder` == 0)
```


#Average grade and age of students with scholarships
```{r}
ggplot(data = scholarship, aes(x = scholarship$`Age at enrollment`, y = scholarship$MeanGrade )) +
  geom_point(color = "blue") +
  labs(title = "Nota media y edad de los alumnos con beca") +
  xlab("Edad al aplicar a la carrera") +
  ylab("Nota media") +
  theme_minimal()
```

  
#Average grade and age of students without scholarship
```{r}
ggplot(data = no_scholarship, aes(x = no_scholarship$`Age at enrollment`, y = no_scholarship$MeanGrade)) +
  geom_point(color = "red") +
  labs(title = "Nota media y edad de los alumnos sin beca") +
  xlab("Edad al aplicar a la carrera") +
  ylab("Nota media") +
  theme_minimal()
```


#Distribution of the average score of students with and without scholarships
```{r}
ggplot(data = data, aes(x = as.factor(data$`Scholarship holder`), y = data$MeanGrade)) +
  geom_boxplot(fill="slateblue", alpha=0.2) +
  labs(title = "Distribución de la nota media de los alumnos con y sin beca") +
  xlab("¿Posee beca?") +
  ylab("Nota media") +
  theme_minimal()
```


#Density plot comparison: MeanGrade between Scholarship holders
```{r}
ggplot(data = data, aes(x = MeanGrade, group = as.factor(data$`Scholarship holder`), fill = as.factor(data$`Scholarship holder`))) +
  geom_density(adjust = 3) +
  facet_wrap(as.factor(data$`Scholarship holder`)) +
  scale_fill_manual(values = c("#0072B2", "#E69F00")) + 
  xlab("Nota media") +
  ylab("Densidad") +
  ggtitle("Densidad de alumnos con nota media dependiendo de si tienen beca")
  theme(
    legend.position = "none",
    panel.spacing = unit(0.1, "lines"),
    axis.ticks.x = element_blank()
  )
```
  
  
#Violin chart
```{r}
ggplot(data = data, aes(x = as.factor(`Scholarship holder`), y = MeanGrade, fill = as.factor(Displaced))) +
  geom_violin(scale = "count") +
  scale_fill_manual(values = c("#0072B2", "#E69F00")) +
  labs(x = "¿Posee beca?", y = "Nota media") +
  ggtitle("Nota media entre alumnos becados y desplazados") +
  theme_bw()
```



# Create ridgeline chart to compare the average grade between scholarship and displaced students
```{r}
ggplot(data = data, aes(x = MeanGrade, y = as.factor(`Scholarship holder`), fill = as.factor(Displaced))) +
  geom_density_ridges(alpha = 0.8, scale = 2, rel_min_height = 0.01, color = "white") +
  scale_fill_manual(values = c("#0072B2", "#E69F00")) +
  labs(x = "Nota media", y = "¿Posee beca?", fill = "¿Alumno desplazado?") +
  ggtitle("Nota media entre alumnos becados y desplazados") +
  theme_ipsum() +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
  )

```


#Test ANOVA
```{r}
modelo <- aov(data$MeanGrade ~ data$`Scholarship holder` + data$Displaced, data=data)
qqnorm(resid(modelo))
qqline(resid(modelo))
anova(modelo)
```


#Test POST HOC
```{r}
modelo <- aov(data$MeanGrade ~ as.factor(data$`Scholarship holder`) + as.factor(data$Displaced), data=data)
tukey <- TukeyHSD(modelo)
tukey
```

